---
title: "ART TREES - TMR Replication Demonstration"
author: "GHG Program Team"
date: "2025-08-08"
output:
  html_document:
    toc: yes
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **1. Introduction: The Goal of This Script**

This R Markdown file is a practical exercise for the ART TREES Verification
Audit Training. Its purpose is to demonstrate **perfect reproducibility**, a
core principle of a successful audit submission.

An auditor must be able to take your data and your code, run it in a documented
environment, and get the exact same result you reported. This script performs a
simple calculation of total carbon stock from a set of hypothetical forest
plots.

**Your task**: Run this script and compare the final output to the value
reported in the training's mock TREES Monitoring Report (TMR), Table 16:
**`1,550`** tCO₂e.

-   **As per ISO 14064-3**, conclusions must be based on verifiable evidence.
    This script and the data within it constitute that evidence.
-   **Data Traceability**: The origin of the final number is transparently
    documented in the code below.

--------------------------------------------------------------------------------

### **2. Loading Required Packages**

This first code step loads the `dplyr` package, which provides tools for data
manipulation. In a real audit, the auditor would ensure this package version
matches the one listed in your `runtime_log.txt` file to ensure a consistent
computational environment.

```{r packages, message=FALSE, warning=FALSE}
# The 'dplyr' package is used for its straightforward data handling functions.
library(dplyr)
```

--------------------------------------------------------------------------------

### **3. Defining the Dataset**

For this demonstration, the data is entered directly into the script. This
ensures that the analysis is self-contained and perfectly reproducible without
relying on external files, which could be misplaced or altered. This represents
our "raw data" for the audit.

Each row represents a forest plot with a unique ID and its measured carbon
stock.

```{r data}
# Create a data frame with hypothetical forest inventory data.
# In a real-world scenario, this data would be loaded from a controlled, versioned file.
# Evidence Source: Corresponds to "Data-Inventory/Field_Plot_Data_2024.csv" in the data repository.
plot_data <- data.frame(
  plot_id = c("A-01", "A-02", "B-01", "B-02", "C-01"),
  carbon_stock_tCO2e = c(300, 350, 250, 400, 250)
)

# Display the data to ensure it has been loaded correctly.
print("Input Data: Hypothetical Forest Plot Carbon Stock")
print(plot_data)
```

--------------------------------------------------------------------------------

### **4. Performing the Calculation**

This is the core of the analysis. We perform a simple sum of the
`carbon_stock_tCO2e` column. The process is intentionally simple to be
transparent. An auditor can easily review this code and confirm that the
methodology is sound and has been implemented correctly.

-   **Accuracy Principle**: The calculation must be free from bias and error. A
    simple sum is the most accurate way to aggregate this data.
-   **Completeness Principle**: All records from the defined dataset are
    included in the calculation.

```{r calculation}
# Calculate the total carbon stock by summing the values in the 'carbon_stock_tCO2e' column.
# This directly replicates the procedure described in the TMR's methodology section.
total_carbon_stock <- plot_data %>%
  summarise(total_stock = sum(carbon_stock_tCO2e))

# Extract the final numerical value for clear reporting.
final_result_tCO2e <- total_carbon_stock$total_stock
```

--------------------------------------------------------------------------------

### **5. Final Result for Audit Verification**

This final chunk displays the single, definitive result of the analysis. This is
the number an auditor would compare against your official TMR submission. A
perfect match demonstrates reproducibility and data integrity.

```{r results}
# Print the final result in a clear, unambiguous format.
cat("--------------------------------------------------\n")
cat("          AUDIT VERIFICATION RESULT\n")
cat("--------------------------------------------------\n")
cat("Total Carbon Stock from All Plots (tCO2e):", final_result_tCO2e, "\n")
cat("--------------------------------------------------\n")
cat("\nThis value should exactly match Table 16 of the TMR.\n")
```

**Conclusion**: If the value `1550` is displayed above, you have successfully
replicated the analysis. This confirms that the data, methods, and environment
are consistent and verifiable, meeting a key requirement for a successful ART
TREES audit.
