---
title: "ART-TREES Data Replication Demo"
subtitle: "Training exercise in version control and reproducible analyses"
author: "Murphy, S"
date: "2025-08-08"
output:
  html_document:
    toc: yes
    theme: united
    df_print: kable
    highlight: pygments
    latex_engine: xelatex
    keep_md: true
    
always_allow_html: TRUE
editor_options: 
  markdown: 
    wrap: 70
---

```{r setup-1}
#| warning: false
#| message: false
#| error: false
#| include: false
#| echo: false
easypackages::packages(
  "animation", "allodb", "BIOMASS", "c2z", "caret", "dataMaid", "DescTools","dplyr",
  "flextable", "FawR", "ForestToolsRS", "ggplot2",
  "janitor", "lattice", "kableExtra", "kernlab", "knitr", 
  "RColorBrewer", "rmarkdown", "readxl", "tibble", "tidymodels", 
  "tidyverse","tinytex", "tune", "useful", "webshot", "webshot2",
  prompt = F
  )
  
knitr::opts_chunk$set(
  echo    = TRUE, 
  message = FALSE, 
  warning = FALSE,
  error   = FALSE, 
  cache   = FALSE,
  comment = NA, 
  tidy.opts = list(width.cutoff = 70)
)

options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
sf::sf_use_s2(use_s2 = FALSE)
```

### **1. Introduction: The Goal of This Exercise**

This R Markdown file demonstrates **perfect reproducibility** in
environmental data analysis, a fundamental principle for any credible
scientific or audit process.

**The Challenge**: An independent reviewer must be able to take your
data and your code, run it in a documented environment, and get the
exact same result you reported.

This script performs a straightforward calculation of total carbon
stock from forest inventory data using publicly available
methodologies.

**Your task**: Run this script and verify that you get the expected
result of **1,550 tCO₂e**. This demonstrates that:

-   Your computational environment is properly configured
-   The analysis is transparent and reproducible\
-   The methodology can be independently verified

**Key Principles Demonstrated**:

-   **Data Traceability**: Every calculation step is documented
-   **Reproducibility**: Results are identical across different users
    and systems
-   **Transparency**: Methods are open and verifiable

----------------------------------------------------------------------

### **2. Loading Required Packages**

This step loads the `dplyr` package for data manipulation. In
practice, all package versions should be documented (see runtime log
at the end of this script) to ensure computational reproducibility.

```{r, class.source = c("numCode", "r", "numberLines"), message=FALSE, warning=FALSE}
# Load required packages for data manipulation
install.packages("easypackages")

easypackages::packages(
  "animation", "allodb", "BIOMASS", "dataMaid", "DescTools","dplyr", "flextble",
  "FawR", "ForestToolsRS", "ggplot2", "janitor", "lattice", "kableExtra", 
  "kernlab", "knitr", "psych", "RColorBrewer", "rmarkdown", "readxl", 
  "tibble", "tidymodels", "tidyverse", "tune", "useful", "webshot", "webshot2",
  prompt = F
  )
  
# Verify package is loaded
cat("All packages loaded successfully")
```

----------------------------------------------------------------------

### **3. Creating the Dataset**

We use a standardized dataset that represents typical forest inventory
measurements. This approach ensures the exercise works consistently
across different environments without requiring external file
dependencies.

**Data Source**: Simulated forest plot measurements following standard
forestry protocols.

```{r, class.source = c("numCode", "r", "numberLines")}
# Create standardized forest inventory dataset
# Each row represents one forest measurement plot
carbon_data <- data.frame(
  plot_id = c("PLOT_001", "PLOT_002", "PLOT_003", "PLOT_004", "PLOT_005"),
  carbon_stock_tCO2e = c(310, 290, 350, 275, 325),
  forest_type = c("Primary", "Secondary", "Primary", "Secondary", "Primary"),
  measurement_year = c(2024, 2024, 2024, 2024, 2024)
)

# Display the input data
cat("Forest Inventory Data:\n")
carbon_data_flex = flextable(head(carbon_data[, 1:4])) |>
  set_header_labels(carbon_data_flex,
  values = list(
    plot_id                 = "Plot ID",
    carbon_stock_tCO2e      = "Carbon Stock (tCO2-e)",
    forest_type             = "Forest Type",
    measurement_year        = "Inventory Date (YYYY)"
  )) |> width(width = 1) |> fit_to_width(max_width = 6.5) |>
  colformat_double(digits = 1, na_str = "N/A")
  #colformat_double(carbon_data_flex, j = "carbon_stock_tCO2e", digits = 1, na_str = "N/A")

carbon_data_flex |> fontsize(size=8,part="all")
cat("\nDataset contains", nrow(carbon_data), "measurement plots\n")
```

```{r, class.source = c("numCode", "r", "numberLines")}

# Create standardized forest inventory dataset
# Each row represents one forest measurement plot
carbon_data <- data.frame(
  plot_id = c("PLOT_001", "PLOT_002", "PLOT_003", "PLOT_004", "PLOT_005"),
  carbon_stock_tCO2e = c(310, 290, 350, 275, 325),
  forest_type = c("Primary", "Secondary", "Primary", "Secondary", "Primary"),
  measurement_year = c(2024, 2024, 2024, 2024, 2024)
)

# Display the input data
cat("Forest Inventory Data:\n")
carbon_data_flex <- flextable(head(carbon_data[, 1:4]))
carbon_data_flex <- set_header_labels(
  carbon_data_flex,
  values = list(
    plot_id             = "Plot ID",
    carbon_stock_tCO2e  = "Carbon Stock (tCO2-e)",
    forest_type         = "Forest Type",
    measurement_year    = "Inventory Date (YYYY)"
    ))

carbon_data_flex <- width(carbon_data_flex, width = 1)
carbon_data_flex <- fit_to_width(carbon_data_flex, max_width = 6.5)

# The 'big.mark' argument has been removed, so no thousand separators will be used.
carbon_data_flex <- colformat_double(
    carbon_data_flex,
    digits = 1,
    na_str = "N/A"
)

carbon_data_flex |> fontsize(size=8,part="all")
cat("\nDataset contains", nrow(carbon_data), "measurement plots\n")
```

----------------------------------------------------------------------
